Identification of high-yielding soybean lines with exceptional seed composition qualities

### 2022-04-20 doc/manuscript/manuscript.Rmd
some small changes to the yield summary table function. Added a table to the utility tables list to convert summary table column names to more easily readible column names
updated manuscript

---
title: Identification of high-yielding soybean lines with exceptional seed composition qualities
author:
  - name: Jay Gillenwater
    affil: 1,2,\ddagger,*
    orcid: 0000-0003-3293-2315
  - name: John Doe
    affil: 2, \dagger, \ddagger
affiliation:
  - num: 1
    address: |
      Muenster University of Applied Sciences - 
      Institute for Infrastructure, Water, Resources, Environment
      Correnstr. 25, 48149 Muenster, Germany
    email: leutnant@fh-muenster.de
  - num: 2
    address: |
      Your department
      Street, City, Country
    email: jhgille2@ncsu.edu
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
secondnote: |
  These authors contributed equally to this work.
correspondence: |
  leutnant@fh-muenster.de; Tel.: +XX-000-00-0000.
journal: Agronomy
type: article
status: submit
bibliography: mybibfile.bib
appendix: appendix.tex
simplesummary: |
  A Simple summary goes here.
abstract: |
  A single paragraph of about 200 words maximum. For research articles, 
  abstracts should give a pertinent overview of the work. We strongly encourage
  authors to use the following style of structured abstracts, but without 
  headings: 1) Background: Place the question addressed in a broad context and
  highlight the purpose of the study; 2) Methods: Describe briefly the main
  methods or treatments applied; 3) Results: Summarize the article's main 
  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
  The abstract should be an objective representation of the article, it must not 
  contain results which are not presented and substantiated in the main text and 
  should not exaggerate the main conclusions.
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
conflictsofinterest: |
  Declare conflicts of interest or state 'The authors declare no conflict of 
  interest.' Authors must identify and declare any personal circumstances or
  interest that may be perceived as inappropriately influencing the
  representation or interpretation of reported research results. Any role of the
  funding sponsors in the design of the study; in the collection, analyses or 
  interpretation of data in the writing of the manuscript, or in the decision to 
  publish the results must be declared in this section. If there is no role, 
  please state 'The founding sponsors had no role in the design of the study; 
  in the collection, analyses, or interpretation of data; in the writing of the 
  manuscript, an in the decision to publish the results'.
sampleavailability: |
  Samples of the compounds ...... are available from the authors.
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: LD 
    long: linear dichroism
output: rticles::mdpi_article
---

@@

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper [@bertrand-krajewski_distribution_1998; @leutnant_stormwater_2016]. And now citing a book reference @gujer_systems_2008. Please use the command [@leutnant_stormwater_2016] for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.

@@

## Population development
In 2018, oil mapping populations 201 and 202 were grown at the Central Crops Research Station in Clayton, NC. These populations consisted of 273 and 237 recombinant inbred lines (RILs) respectively. Several agronomic traits were scored in the field for each population. These included height, lodging, maturity date, and a composite agronomic score. Lodging was scored on a scale of 1-5 where 5indicates that all plants in a plot are on the ground, and a score of 1 indicates that all plants are erect[Feh87]. The agronomic score aimed to capture other traits of value such as visual estimation of pod load and plot uniformity to provide a general score of a line’s agronomic desirability. Agronomic score was recorded on a scale of 1-5 aswell, with 1 identifying the best lines of a population, and 5 the worst. Maturity was recorded at the R8 maturity date. Height was measured in centimeters from the soil to the top of the plant. Following harvest, yield, seed weight, protein, and oil content were measured after seed was air dried to approximately 7% moisture content in a greenhouse. Protein and oil contents were measured on a dry basis using a Perten DA 7250 NIR®instrument. Yield and seed weight were measured after seed had been sifted and cleaned of debris and cracked seed. To select lines for the 2019 growing season, lines with abnormally low bulk weights or extreme maturity dates from 2018 were first removed from consideration. Two yield trials were then developed for each mapping population. The maturity data of RILs were considered when forming tests such that the lines of each test would have a maturity data range approximately half that of the total mapping population from which it was derived. RILs were selected for each test which were also representative of the distribution of seed protein and seed oil traits for each population. Eighty unique lines were selected from each population which satisfied these criteria, and each yield test was comprised of 40 RILs. Three high-yielding check cultivars and the two parents of the respective population were also included in each test. Yield check cultivars Dunphy, Osage, and Roy were used in tests 1 and 2, while Dunphy, Dilday, and N.C. Raleigh were used for tests 3 and 4. These lines were selected to represent the estimated maturities of the RILs in each test. The parents for tests 1 and 2 were cultivars LMN09-119 and N09-09, and the parents for tests 3 and 4 were LMN09-19 and N13-47.

These four tests were grown in two locations in 2019: the Tidewater Research Station in Plymouth, NC (PLY) and the Caswell Research Farm in Kinston, NC (CAS). The same data was collected for each test in this season that was collected in the previous season. Selections were done using this data by first identifying the lines with a yield within or above a least significant difference (LSD) of the average yield of the checks for each test. Further selection was done using the seed composition traits by identifying the thirty lines with the highest protein + oil content.

These thirty lines were then grouped into two new tests of 15 RILs each which were then grown in both the 2020 and 2021 seasons. Thesew tests were grown in CLA and CAS in 2020 and CAS and PLY in 2021.

>>>>

---
title: Identification of high-yielding soybean lines with exceptional seed composition qualities
author:
  - name: Jay Gillenwater
    affil: 1,\ddagger,*
    orcid: 0000-0002-3477-6291
affiliation:
  - num: 1
    address: |
      Department of Crop and Soil Sciences, NCSU, Raleigh, NC, USA
    email: jhgille2@ncsu.edu
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
secondnote: |
  These authors contributed equally to this work.
correspondence: |
  leutnant@fh-muenster.de; Tel.: +XX-000-00-0000.
journal: Agronomy
type: article
status: submit
bibliography: mybibfile.bib
appendix: appendix.tex
simplesummary: |
  A Simple summary goes here.
abstract: |
  A single paragraph of about 200 words maximum. For research articles, 
  abstracts should give a pertinent overview of the work. We strongly encourage
  authors to use the following style of structured abstracts, but without 
  headings: 1) Background: Place the question addressed in a broad context and
  highlight the purpose of the study; 2) Methods: Describe briefly the main
  methods or treatments applied; 3) Results: Summarize the article's main 
  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
  The abstract should be an objective representation of the article, it must not 
  contain results which are not presented and substantiated in the main text and 
  should not exaggerate the main conclusions.
keywords: |
  yield; protein; oil;soybean; protein meal
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
conflictsofinterest: |
  Declare conflicts of interest or state 'The authors declare no conflict of 
  interest.' Authors must identify and declare any personal circumstances or
  interest that may be perceived as inappropriately influencing the
  representation or interpretation of reported research results. Any role of the
  funding sponsors in the design of the study; in the collection, analyses or 
  interpretation of data in the writing of the manuscript, or in the decision to 
  publish the results must be declared in this section. If there is no role, 
  please state 'The founding sponsors had no role in the design of the study; 
  in the collection, analyses, or interpretation of data; in the writing of the 
  manuscript, an in the decision to publish the results'.
sampleavailability: |
  Samples of the compounds ...... are available from the authors.
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: LD 
    long: linear dichroism
output: rticles::mdpi_article
---

@@

Seed yield, oil, and protein are all valuable traits in a soybean variety, however breeding lines which have both high yield and protein has been difficult to develop due to the negative correlation between the two traits[@burton1987quantitative; @ProtOilCorr; @ProtOilCorr_new]. While considerable efforts have been made to identify loci which control these seed quality traits so that MAS breeding strategies can be utilized for their improvement, to date the applications of such markers have been few. This is largely due to the lack of markers which are uniquely associated with one trait, and are also stable across genetic and environmental backgrounds. While there is still reason to continue this genetic research, it is important that breeders take every opportunity to identify lines with both high yield, and seed composition traits like oil and protein content so that new varieties can be released.

Soybean lines typically contain about 20\% oil and 40\% protein content on a dry weight basis[@ProteinGenomics]. The market for soybean meal requires 47.5\% protein content in the meal, which corresponds to approximately 41.5\% protein content on a dry weight basis[@ProteinGenomics]. Oil and protein content are two of the most important seed composition traits in soybean so if one is decreased, the other should be correspondingly increased to account for the loss in value. The inverse correlation between protein and oil contents is well known and is suspected to be due at least partially to the action of pleiotropic genes and competing metabolic pathways which control the expression of each trait[@SeedCompositionGenomics].

Despite the difficulty in simultaneously breeding for all three of these traits, releases of varieties with elevated protein contents and only moderately reduced yield has shown that it is not impossible. The high protein germplasm lines R05-1415 and R05-1772 were released recently and contain 46.9\% and 46.1\% protein content with while still producing yields 94\% and 91\% of that of the high yielding 5002T cultivar[@HiProtHiYield02]. Lines TN03-350 and TN04-5321 contain 43.9\% and 43.1\% protein content while having superior or comparable performance to yield checks[@HiProtHiYield04]. The Prolina cultivar has a protein content of 46.1\% with a yield only 13\% reduced from the Centennial check[@HiProtHiYield01]. The Highpro1 cultivar was released in 2016 and has a yield which is greater than or equal to 97\% of that of the highest yielding check cultivar, IA3023 with a protein content of 40.1\%[@HighPro1]. Cultivars produced through conventional breeding techniques such as these have shown that it is possible to identify lines with both high seed protein and seed yield. Efforts to find these lines should be continued to provide growers and breeders with additional high value cultivars, and germplasm which can be further used to improve protein and yield traits.

To meet this goal, two recombinant inbred line (RIL) oil mapping populations were screened for lines which showed promising combinations of yield and seed composition traits. Successive rounds of selection were conducted to identify and characterize lines with high values for yield as well as protein and oil composition were performed between 2018 and 2021 to identify soybean lines with elevated yield as well as the valuable seed composition traits when compared with existing check cultivars.

@@

## Population development
In 2018, oil mapping populations 201 and 202 were grown as plant rows at the Central Crops Research Station in Clayton, NC. These populations consisted of 273 and 237 recombinant inbred lines (RILs) respectively. Several agronomic traits were scored in the field for each population.

The agronomic traits recorded in the field were height, lodging, maturity date, and a composite agronomic score. Lodging was scored on a scale of 1-5 where 5 indicates that all plants in a plot are on the ground, and a score of 1 indicates that all plants are erect[@fehr1987soybeans]. The agronomic score aimed to capture other traits of value such as visual estimation of pod load and plot uniformity to provide a general score of a line’s agronomic desirability. Agronomic score was recorded on a scale of 1-5 as well, with 1 identifying the best lines of a population, and 5 the worst. Maturity was recorded at the R8 maturity date and was recorded as the number of days after September 1. Height was measured in inches from the soil to the top of the plant.

Following harvest, yield, seed weight, protein, and oil content were measured after seed was air dried to approximately 7% moisture content in a greenhouse. Protein and oil contents were measured on a dry basis using a Perten DA 7250 NIR®instrument. Yield and seed weight were measured after seed had been sifted and cleaned of debris and cracked seed.

To select lines for the 2019 growing season, lines with abnormally low bulk weights or extreme maturity dates from 2018 were first removed from consideration. Two yield trials were then developed for each mapping population. The maturity dates of RILs were considered when forming tests such that the lines of each test would have a maturity date range approximately half that of the total mapping population from which it was derived. RILs were selected for each test which were also representative of the distribution of seed protein and seed oil traits for each population.

Eighty unique lines were selected from each population which satisfied these criteria, and each yield test was comprised of 40 RILs. Three high-yielding check cultivars and the two parents of the respective population were also included in each test. Yield check cultivars Dunphy, Osage[@Osage], and Roy were used in tests 1 and 2, while Dunphy, Dilday, and NC-Raleigh[@NCRaleighregistration] were used for tests 3 and 4. These lines were selected to represent the estimated maturities of the RILs in each test. The parents for tests 1 and 2 were cultivars LMN09-119 and N09-09, and the parents for tests 3 and 4 were LMN09-19 and N13-47.

These four tests were grown in two locations in 2019: the Tidewater Research Station in Plymouth, NC (PLY) and the Caswell Research Farm in Kinston, NC (CAS). The same data was collected for each test in this season that was collected in the previous season.

Using the data collected from the 2019 season season, further selections were done to identify high-yielding lines from the four tests. This was done by identifying the RILs with a yield within or above a least significant difference (LSD) of the average yield of the checks for each test. Further selection was done using the seed composition traits by identifying the thirty RILs with the highest protein + oil content on a dry basis from among the RILs which had passed the yield selection threshold.

These thirty lines were then grouped into two new tests of 15 RILs each based on maturity date. These two new tests are named Test 1 and Test 2. Yield check cultivars were again assigned to each test to match the maturity dates of the RILs that were in each test. Cultivars Dunphy, DIlday, and NC-Raleigh were used as checks in Test 1 and Dunphy, Ellis, N10-697, and Osage were used as checks in Test 2.

These two tests were grown in both the 2020 and 2021 seasons. These tests were grown in CLA and CAS in 2020 and CAS and PLY in 2021. The same phenotypes were evaluated for each genotype in the 2020 and 2021 seasons using the same methodology that was employed in the 2019 season.

## Statistical Analysis
Phenotypic traits were analysed with a linear model with the form:

$$y_{ijk} = \mu + L_i + B(L_i) + G_k + GL_{ik} + \epsilon_{ijk}$$

Where $y_{ijk}$ is the phenotypic measurement for rep$j$ of genotype $k$ in environment $i$, $L_i$ is the effect of location $i$, $G_k$ is the effect of genotype $G$, $GL_{ik}$ is the interaction effect of location $L$ and genotype $G$, and $\epsilon_{ijk}$ is the measurement error.

Models were fit using the lm function in R[@rcoreteamLanguageEnvironmentStatistical2022], and analysis of variance (ANOVA) performed using the anova function. Least-square means (LS Means) for each genotype and trait were calculated using the above model using the emmeans package[@lenthEmmeansEstimatedMarginal2022] in R. Least-significant difference values for each trait were calculated using the LSD.test function from the agricolae[@de2014agricolae] package. Pearson correlation coefficients between each phenotype were calculated with the cor function in R.

Pearson correlation is calculated for each pair of traits as:

$$r = \frac{\sum{(x-m_x)(y-m_y)}}{\sqrt{\sum{(x-m_x)^2}\sum{(y-m_y)^2}}}$$

Where $x$ and $y$ are measurements of the two phenotypes, $m_x$ and $m_y$ are the means of each phenotype, and $r$ is the correlation coefficient.

### 2022-04-20 doc/manuscript/manuscript.tex
updated manuscript

% Authors, for the paper (add full first names)
\Author{Jay
Gillenwater$^{1,2,\ddagger,*}$\href{https://orcid.org/0000-0003-3293-2315}{\orcidicon}, John
Doe$^{2, \dagger, \ddagger}$}

% Authors, for metadata in PDF
\AuthorNames{Jay Gillenwater, John Doe}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Muenster University of Applied Sciences - Institute for
Infrastructure, Water, Resources, Environment Correnstr. 25, 48149
Muenster,
Germany; \href{mailto:leutnant@fh-muenster.de}{\nolinkurl{leutnant@fh-muenster.de}}\\
$^{2}$ \quad Your department Street, City,
Country; \href{mailto:jhgille2@ncsu.edu}{\nolinkurl{jhgille2@ncsu.edu}}\\
}
% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:leutnant@fh-muenster.de}{\nolinkurl{leutnant@fh-muenster.de}};
Tel.: +XX-000-00-0000.}

@@

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent
keywords specific to the article, yet reasonably common within the
subject discipline.).}

@@

The introduction should briefly place the study in a broad context and
highlight why it is important. It should define the purpose of the work
and its significance. The current state of the research field should be
reviewed carefully and key publications cited. Please highlight
controversial and diverging hypotheses when necessary. Finally, briefly
mention the main aim of the work and highlight the principal
conclusions. As far as possible, please keep the introduction
comprehensible to scientists outside your particular field of research.
Citing a journal paper
\citep{bertrand-krajewski_distribution_1998, leutnant_stormwater_2016}.
And now citing a book reference \citet{gujer_systems_2008}. Please use
the command \citep{leutnant_stormwater_2016} for the following MDPI
journals, which use author-date citation: Administrative Sciences, Arts,
Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws,
Religions, Risks, Social Sciences.

@@

In 2018, oil mapping populations 201 and 202 were grown at the Central
Crops Research Station in Clayton, NC. These populations consisted of
273 and 237 recombinant inbred lines (RILs) respectively. Several
agronomic traits were scored in the field for each population. These
included height, lodging, maturity date, and a composite agronomic
score. Lodging was scored on a scale of 1-5 where 5indicates that all
plants in a plot are on the ground, and a score of 1 indicates that all
plants are erect{[}Feh87{]}. The agronomic score aimed to capture other
traits of value such as visual estimation of pod load and plot
uniformity to provide a general score of a line's agronomic
desirability. Agronomic score was recorded on a scale of 1-5 aswell,
with 1 identifying the best lines of a population, and 5 the worst.
Maturity was recorded at the R8 maturity date. Height was measured in
centimeters from the soil to the top of the plant. Following harvest,
yield, seed weight, protein, and oil content were measured after seed
was air dried to approximately 7\% moisture content in a greenhouse.
Protein and oil contents were measured on a dry basis using a Perten DA
7250 NIR®instrument. Yield and seed weight were measured after seed had
been sifted and cleaned of debris and cracked seed. To select lines for
the 2019 growing season, lines with abnormally low bulk weights or
extreme maturity dates from 2018 were first removed from consideration.
Two yield trials were then developed for each mapping population. The
maturity data of RILs were considered when forming tests such that the
lines of each test would have a maturity data range approximately half
that of the total mapping population from which it was derived. RILs
were selected for each test which were also representative of the
distribution of seed protein and seed oil traits for each population.
Eighty unique lines were selected from each population which satisfied
these criteria, and each yield test was comprised of 40 RILs. Three
high-yielding check cultivars and the two parents of the respective
population were also included in each test. Yield check cultivars
Dunphy, Osage, and Roy were used in tests 1 and 2, while Dunphy, Dilday,
and N.C. Raleigh were used for tests 3 and 4. These lines were selected
to represent the estimated maturities of the RILs in each test. The
parents for tests 1 and 2 were cultivars LMN09-119 and N09-09, and the
parents for tests 3 and 4 were LMN09-19 and N13-47.

These four tests were grown in two locations in 2019: the Tidewater
Research Station in Plymouth, NC (PLY) and the Caswell Research Farm in
Kinston, NC (CAS). The same data was collected for each test in this
season that was collected in the previous season. Selections were done
using this data by first identifying the lines with a yield within or
above a least significant difference (LSD) of the average yield of the
checks for each test. Further selection was done using the seed
composition traits by identifying the thirty lines with the highest
protein + oil content.

These thirty lines were then grouped into two new tests of 15 RILs each
which were then grown in both the 2020 and 2021 seasons. Thesew tests
were grown in CLA and CAS in 2020 and CAS and PLY in 2021.

>>>>

% Authors, for the paper (add full first names)
\Author{Jay
Gillenwater$^{1,\ddagger,*}$\href{https://orcid.org/0000-0002-3477-6291}{\orcidicon}}

% Authors, for metadata in PDF
\AuthorNames{Jay Gillenwater}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Department of Crop and Soil Sciences, NCSU, Raleigh, NC,
USA; \href{mailto:jhgille2@ncsu.edu}{\nolinkurl{jhgille2@ncsu.edu}}\\
}
% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:leutnant@fh-muenster.de}{\nolinkurl{leutnant@fh-muenster.de}};
Tel.: +XX-000-00-0000.}

@@

% Keywords
\keyword{yield; protein; oil;soybean; protein meal}

@@

Seed yield, oil, and protein are all valuable traits in a soybean
variety, however breeding lines which have both high yield and protein
has been difficult to develop due to the negative correlation between
the two
traits\citep{burton1987quantitative, ProtOilCorr, ProtOilCorr_new}.
While considerable efforts have been made to identify loci which control
these seed quality traits so that MAS breeding strategies can be
utilized for their improvement, to date the applications of such markers
have been few. This is largely due to the lack of markers which are
uniquely associated with one trait, and are also stable across genetic
and environmental backgrounds. While there is still reason to continue
this genetic research, it is important that breeders take every
opportunity to identify lines with both high yield, and seed composition
traits like oil and protein content so that new varieties can be
released.

Soybean lines typically contain about 20\% oil and 40\% protein content
on a dry weight basis\citep{ProteinGenomics}. The market for soybean
meal requires 47.5\% protein content in the meal, which corresponds to
approximately 41.5\% protein content on a dry weight
basis\citep{ProteinGenomics}. Oil and protein content are two of the
most important seed composition traits in soybean so if one is
decreased, the other should be correspondingly increased to account for
the loss in value. The inverse correlation between protein and oil
contents is well known and is suspected to be due at least partially to
the action of pleiotropic genes and competing metabolic pathways which
control the expression of each trait\citep{SeedCompositionGenomics}.

Despite the difficulty in simultaneously breeding for all three of these
traits, releases of varieties with elevated protein contents and only
moderately reduced yield has shown that it is not impossible. The high
protein germplasm lines R05-1415 and R05-1772 were released recently and
contain 46.9\% and 46.1\% protein content with while still producing
yields 94\% and 91\% of that of the high yielding 5002T
cultivar\citep{HiProtHiYield02}. Lines TN03-350 and TN04-5321 contain
43.9\% and 43.1\% protein content while having superior or comparable
performance to yield checks\citep{HiProtHiYield04}. The Prolina cultivar
has a protein content of 46.1\% with a yield only 13\% reduced from the
Centennial check\citep{HiProtHiYield01}. The Highpro1 cultivar was
released in 2016 and has a yield which is greater than or equal to 97\%
of that of the highest yielding check cultivar, IA3023 with a protein
content of 40.1\%\citep{HighPro1}. Cultivars produced through
conventional breeding techniques such as these have shown that it is
possible to identify lines with both high seed protein and seed yield.
Efforts to find these lines should be continued to provide growers and
breeders with additional high value cultivars, and germplasm which can
be further used to improve protein and yield traits.

To meet this goal, two recombinant inbred line (RIL) oil mapping
populations were screened for lines which showed promising combinations
of yield and seed composition traits. Successive rounds of selection
were conducted to identify and characterize lines with high values for
yield as well as protein and oil composition were performed between 2018
and 2021 to identify soybean lines with elevated yield as well as the
valuable seed composition traits when compared with existing check
cultivars.

@@

In 2018, oil mapping populations 201 and 202 were grown as plant rows at
the Central Crops Research Station in Clayton, NC. These populations
consisted of 273 and 237 recombinant inbred lines (RILs) respectively.
Several agronomic traits were scored in the field for each population.

The agronomic traits recorded in the field were height, lodging,
maturity date, and a composite agronomic score. Lodging was scored on a
scale of 1-5 where 5 indicates that all plants in a plot are on the
ground, and a score of 1 indicates that all plants are
erect\citep{fehr1987soybeans}. The agronomic score aimed to capture
other traits of value such as visual estimation of pod load and plot
uniformity to provide a general score of a line's agronomic
desirability. Agronomic score was recorded on a scale of 1-5 as well,
with 1 identifying the best lines of a population, and 5 the worst.
Maturity was recorded at the R8 maturity date and was recorded as the
number of days after September 1. Height was measured in inches from the
soil to the top of the plant.

Following harvest, yield, seed weight, protein, and oil content were
measured after seed was air dried to approximately 7\% moisture content
in a greenhouse. Protein and oil contents were measured on a dry basis
using a Perten DA 7250 NIR®instrument. Yield and seed weight were
measured after seed had been sifted and cleaned of debris and cracked
seed.

To select lines for the 2019 growing season, lines with abnormally low
bulk weights or extreme maturity dates from 2018 were first removed from
consideration. Two yield trials were then developed for each mapping
population. The maturity dates of RILs were considered when forming
tests such that the lines of each test would have a maturity date range
approximately half that of the total mapping population from which it
was derived. RILs were selected for each test which were also
representative of the distribution of seed protein and seed oil traits
for each population.

Eighty unique lines were selected from each population which satisfied
these criteria, and each yield test was comprised of 40 RILs. Three
high-yielding check cultivars and the two parents of the respective
population were also included in each test. Yield check cultivars
Dunphy, Osage\citep{Osage}, and Roy were used in tests 1 and 2, while
Dunphy, Dilday, and N.C. Raleigh were used for tests 3 and 4. These
lines were selected to represent the estimated maturities of the RILs in
each test. The parents for tests 1 and 2 were cultivars LMN09-119 and
N09-09, and the parents for tests 3 and 4 were LMN09-19 and N13-47.

These four tests were grown in two locations in 2019: the Tidewater
Research Station in Plymouth, NC (PLY) and the Caswell Research Farm in
Kinston, NC (CAS). The same data was collected for each test in this
season that was collected in the previous season.

Using the data follected from this season, further selections were done
to identify high-yielding lines from the four tests. This was done by
identifying the RILs with a yield within or above a least significant
difference (LSD) of the average yield of the checks for each test.
Further selection was done using the seed composition traits by
identifying the thirty RILs with the highest protein + oil content on a
dry basis from among the RILs which had passed the yield selection
threshold.

These thirty lines were then grouped into two new tests of 15 RILs each
on the basis of maturity date. These two new tests are named Test 1 and
Test 2. Yield check cultivars were again assigned to each test to match
the maturity dates of the RILs that were in each test. Cultivars Dunphy,
DIlday, and NC-Raleigh were used as checks in Test 1 and Dunphy, Ellis,
N10-697, and Osage were used as checks in Test 2.

These two tests were grown in both the 2020 and 2021 seasons. These
tests were grown in CLA and CAS in 2020 and CAS and PLY in 2021. The
same phenotypes were evaluated for each genotype in the 2020 and 2021
seasons using the same methodology that was employed in the 2019 season.

\hypertarget{statistical-analysis}{%
\subsection{Statistical Analysis}\label{statistical-analysis}}

Phenotypic traits were analysed with a linear model with the form:

\[y_{ijk} = \mu + L_i + B(L_i) + G_k + GL_{ik} + \epsilon_{ijk}\]

Where \(y_{ijk}\) is the phenotypic measurement for rep\(j\) of genotype
\(k\) in environment \(i\), \(L_i\) is the effect of location \(i\),
\(G_k\) is the effect of genotype \(G\), \(GL_{ik}\) is the interaction
effect of location \(L\) and genotype \(G\), and \(\epsilon_{ijk}\) is
the measurement error.

Models were fit using the lm function in
R\citep{rcoreteamLanguageEnvironmentStatistical2022}, and analysis of
variance (ANOVA) performed using the anova function. Least-square means
(LS Means) for each genotype and trait were calculated using the above
model using the emmeans package\citep{lenthEmmeansEstimatedMarginal2022}
in R. Least-significant difference values for each trait were calculated
using the LSD.test function from the agricolae\citep{de2014agricolae}
package. Pearson correlation coefficients between each phenotype were
calculated with the cor function in R.

Pearson correlation is calculated for each pair of traits as:

\[r = \frac{\sum{(x-m_x)(y-m_y)}}{\sqrt{\sum{(x-m_x)^2}\sum{(y-m_y)^2}}}\]

Where \(x\) and \(y\) are measurements of the two phenotypes, \(m_x\)
and \(m_y\) are the means of each phenotype, and \(r\) is the
correlation coefficient.
